---
title: "R Notebook"
output: html_notebook
---
```{r}
library(psych)
library(corrplot)
library(dplyr)
library(tidyverse)
library(stringr)
library(base)
library(sf)
library(scales)
library(vars)
library(lattice)
library(urca)
library(ggplot2)
library(MASS)
```

# read file
```{r}
data=read.csv('new_data.csv')
```

# Data exploration
```{r}
describe(data)
```
There are 12 variables in the data with 485 observations. Including 1 date variable, 2 response variables, and 9 predictors. All variables are numeric type except variale 'date'. 
1. date
2. new_cases: the daily new confirmed COVID cases in Netherlands
3. new_deaths: the daily new deaths of COVID patients in Netherlands
4. ICU_patients: the COVID patients admitted to ICU per day in hospital
5. Hosp_patients: the COVID patients admitted to hosptial per day
6. stringency_index:
7. temp: daily temperature
8. humidity: daily humidity
9. precip: the amount of rains per day
10. windspeed: daily windspeed
11. transit_stations:
12: residential:

问题： transit 都是负数，需不需要调整从0开始 判断人们乘坐交通的变化

## checking missingness
```{r}
sum(is.na(data))
sum(is.na(data$icu_patients))
sum(is.na(data$hosp_patients))
```
There are 18 missing values in the data, locate in ICU patients and Hospital patients columns. (The government website does not have COVID records for some days during 2021- 2022)

## correlations 
```{r}
pairs(data[, 2:6])

cor(data[, 2:12])

corrplot(cor(data[, 2:12]), method = "circle",
        title = "method = 'circle'") 
```

For hospital admission
1.hospital admission and humidity, hospital admission and precipitation have small correlation less than 0.1
2.hospital admission and new deaths, hospital admission and resident are highly positive correlated 
3.hospital admission and temperature, hospital admission and transit are highly negative correlated

for hospital admission: fit var model including all variables, comparing to the var model without humidity, precipitation, and wind speed.



For ICU admission
1.ICU admission and humidity, ICU admission and precipitation, ICU and wind speed have small correlation less than 0.1.
2.ICU admission and new deaths, ICU admission and stringency, ICU admission and resident are highly positive correlated 
3.ICU admission and temperature, ICU admission and transit are highly negative correlated

for ICU admission: fit var model including all variables, comparing to the var model without humidity, precipitation and wind speed.


new deaths highly positive correlated to stringency_index, and residential. Negative correlated to temperature and public transition.

strigency_index increases, residential increases, transition decreases. (temperature has opposite relation)

transit and residential as highly negative correlated.

# Plots
```{r}
plot(data$new_cases) # 2022-02-07 has 380399 cases, (outlier, remove and graph again) volatity
plot(data$new_cases[-403])
plot(data$new_deaths)
# more deaths at the beginning 
plot(data$icu_patients)
plot(data$hosp_patients)
plot(data$stringency_index)
plot(data$temp)
plot(data$humidity)
plot(data$precip)
plot(data$windspeed)
plot(data$transit_stations)
plot(data$residential)
# ICU, hosp, stringency, residential trends are similar, transition has opposite trend.

```

stringency 会导致lockdown， residential的变化和stringency相似， 从而影响hosp admission。



summer low COVID, winter high COVID
湿度看着和hosp admission， icu走势有点像， 风速和降雨量看不出来。

