---
title: "Time series for hospital patients"
---
```{r,include=FALSE}
library(plyr)
library(imputeTS)
library(dplyr)
library(tidyverse)
library(forcats)
library(stringr)
library(base)
library(sf)
library(scales)
library(lubridate)
library(aTSA)
library(tseries)
library(forecast)
library(CADFtest)
library(vars)
library(urca)
library(lattice)
library(ggplot2)
library(MASS)
library(tidymodels)
library(rugarch)
library(garchmodels)
library(rmgarch)
library(modeltime)
library(timetk)
```

# read data files
```{r}
weekly1=read.csv('data_weekly_generate.csv')

#weekly1$school_open=ifelse(weekly1$school==1, 1, 0)
#weekly1$school_close=ifelse(weekly1$school==1, 0, 1)
```


```{r}
```

# split the data into train and test sets
we use 80% of data as training dataset, 20% of data as test dataset
```{r}
cut_off=0.8*length(weekly1$new_cases)
var_train=weekly1[1:130,]
var_test=weekly1[131:length(weekly1$new_cases),]

#arima_train=ts_df[1:cut_off,]
#arima_test=ts_df[cut_off:length(icu.ts),]
```

# Hospital admission VAR model
The VAR model selects lag=1, which means the values of factors in week t-1 have impact on value in week t of the response.
```{r}

var_hosp=function(df, p, n){
  fit_var<-VAR(df,type="const",p=p) 
  print(AIC(fit_var))
  print(summary(fit_var))
  print(causality(fit_var, cause=colnames(df)[2:length(colnames(df))]))
  pred_var <- predict(fit_var, n.ahead = n)
  hosp_forecast<-ts(pred_var$fcst$hosp_patients[,1],start=2022)
  return(hosp_forecast)
}
```

```{r}
################### full model 
m1=var_train[, c(5,2, 3,6,7,8,9,10,11,12, 13, 14, 15)]

VARselect(m1,type="const", lag.max=10) 
hosp_fore1=var_hosp(m1,6, 7)
hosp_fore1
# there is causality on hospital admission

#when no cases included, prediction ok
# p=6 aic= 5200.115
# p=8 based on VAR aic=4271.254 

#when all variables are included, not too accurate
# p=8 based on VAR, aic=5055.777
# p=6 aic=7311.223

# cases can be removed

```

#some model
#hospital t-1 pos, t-3 neg
#temperature t-1, t-2, t-3 neg
#cases_diff t-1 pos, t-4, neg
#death t-1 neg, t-2 pos, t-3 pos
#stringency t-1 pos, 
#precip t-1 pos
#school t-1 neg, t-2 pos, t-3 neg 
#Residual standard error: 42.62 on 20 degrees of freedom
#Multiple R-Squared: 0.9984,	Adjusted R-squared: 0.996 
#F-statistic: 414.9 on 30 and 20 DF,  p-value: < 2.2e-16 

```{r}
######################## remove humidity, precip, wind, lag=5 week
m2<-var_train[, c(5, 2,3,6,7,8,9,11,12,13,14,15)]
VARselect(m2,type="const", lag.max=10) 
hosp_fore2=var_hosp(m2, 3, 4)
hosp_fore2

# only remove humidity 8
# aic 5652.337, p=8 no
# aic 3759.935, p=9 no
# aic 7240.124, p=3 ok
# aic 7140.25, p=4 nah

# only remove precip 9
# aic 3962.752, p=9 no based on var
# aic 7563.258, p=2 no
# aic 7491.439, p=3 ok
# aic 7367.757, p=4 no

#only remove wind 10
# aic=2810.296, p=9 based on var, no
# aic 6899.918, p=5, ok..
# aic 7162.161, p=3, ok

```

```{r}
######################## remove humidity, precip, wind, cases
m3<-var_train[, c(5,3,6,7,8,9,11,12,13,14,15)]
VARselect(m3,type="const", lag.max=10) 
hosp_fore3=var_hosp(m3, 2, 4)
hosp_fore3
# remove cases, hum, precip, wind together
#aic 7004.437, p=1 based on VAR, #no accurate
#aic 3465.668, p=3
#aic 3476.237, p=2
#aic 3452.608, p=4
#aic 2547.823, p=10 based on aic var, not too accurate

#only remove cases and humidity
# aic  3500.217, p=9 var, ok
# aic 4397.985, p=7 no
# aic 4840, p=3 no

#only remove cases and precip
# aic 3654.069, p=9 var, ok..
# aic 4919.492, p=5, no
# aic 4805.402, p=6, ok
# aic 5089.185, p=3, no

#only remove cases and wind
# aic 3083.889, p=9, var, ok..
# aic 4265.628, p=7, no
# aic 4769.358, p=3, no
# aic 4780.211, p=2, no

```


# computing model errors
```{r}
mape=function(test, forecast, n){
  mape1=mean(abs(((test-forecast)/test)*100))
  return(mape1)
}

mse=function(test, forecast, n){
  mse1=mean((test-forecast)^2)
  return(mse1)
}

rmse=function(test, forecast, n){
  rmse1=sqrt(mean((test-forecast)^2))
  return(rmse1)
}

mae=function(test, forecast, n){
  mae1=mean(abs((test-forecast)))
  return(mae1)
}
```

# comparing models for hospital admission
```{r}
mse1=mse(var_test$hosp_patients[1:4], hosp_fore1, 4)
mse2=mse(var_test$hosp_patients[1:4], hosp_fore2, 4)
mse3=mse(var_test$hosp_patients[1:4], hosp_fore3, 4)
#mse5=mse(var_test$hosp_diff[1:4], hosp_fore5, 4)

rmse1=rmse(var_test$hosp_patients[1:4], hosp_fore1, 4)
rmse2=rmse(var_test$hosp_patients[1:4], hosp_fore2, 4)
rmse3=rmse(var_test$hosp_patients[1:4], hosp_fore3, 4)

mae1=mae(var_test$hosp_patients[1:4], hosp_fore1, 4)
mae2=mae(var_test$hosp_patients[1:4], hosp_fore2, 4)
mae3=mae(var_test$hosp_patients[1:4], hosp_fore3, 4)

mape1=mape(var_test$hosp_patients[1:4], hosp_fore1, 4)
mape2=mape(var_test$hosp_patients[1:4], hosp_fore2, 4)
mape3=mape(var_test$hosp_patients[1:4], hosp_fore3, 4)

#AIC() need AIC value
c(mape1,mape2, mape3)
c(rmse1,rmse2, rmse3)
c(mae1,mae2, mae3)
```

#not useful
```{r}
h=4
s=120
df=weekly1[, c(5,2,3,6,7,8,9,10,11,12,13,14,15)]

error=c()
prediction_vector <- vector(mode="numeric",length=length(df$hosp_patients))
for(k in s:(length(df$hosp_patients)-h)){
  var_df=df[1:k,]
  result=VARselect(var_df,type="const", lag.max=10) 
  print(result)
  hosp_fore=VAR(var_df, p=8,type='const')
  #print(summary(hosp_fore))
  predict_out=predict(hosp_fore, n.head=h)
  predict=predict_out[[1]][1]$hosp_patients[h]
  error=c(error, df$hosp_patients[k+h]-predict)
  prediction_vector[k+h]=predict
}

prediction_vector 

```

